---
title: "Simulations of estimating causal effects using undersmoothed HAL"
author: "Seraphina Shi"
date: "2023-05-10"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_lib, include = FALSE, warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(data.table)
library(dplyr)
library(tidyr)
library(foreach)

library(stringr)
library(glmnet)

library(origami)
library(hal9001)
library(tictoc)

library(R.utils)

library(pROC)

library(ggplot2)
library(cowplot)
library(gridExtra)
library(grid)
```

```{r setup, include = FALSE} 
plotFolder <- here("results","images", "v4")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=paste0(plotFolder, "/"),
  cache.path=".cache/",
  duplicate.label="allow"
)

source(here("scripts", "scripts_v4", "1_simu_functions_hal9001.R"))
source(here("scripts", "scripts_v4", "1_simu_functions.R"))
source(here("scripts", "scripts_v4", "1_visual_functions.R"))
```

# Simulation 1
Data structure:  $O = (W, A, Y)$

 * U - exogenous variables
 * W - baseline covariate that is a measure of body condition
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-5 + W + 2.25A -0.5WA)]$

Outcome of interest: $E_0[Y|a,W]$

```{r check_sys1}
generate_data_1 <- function(n, a=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 1)
  U_A <- rnorm(n, 0, 2)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  
  Y <- as.numeric(U_Y < plogis(-5 + W + 2.25*A - 0.5 * W * A ))
  
  # data frame
  O <- data.frame(W, A, Y)
  return(O)
}


DGS <- generate_data_1
sn = 1
```

```{r child = '2_1_Simu_results_template.Rmd'}
```


# Simulation 2
Data structure:  $O = (W, A, Y)$

 * U - exogenous variables
 * W - baseline covariate that is a measure of body condition
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-10 + 2W + 5sin(A^{1.5}) + 2WA)]$
  
Outcome of interest: $E_0[Y|a,W]$

```{r check_sys2}
generate_data_2 <- function(n, a=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 1)
  U_A <- rnorm(n, 0, 2)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  
  Y <- as.numeric(U_Y < plogis(-10 + 2*W + 5*sin(A^1.5) + 2 * W * A ))
  
  # data frame
  O <- data.frame(W, A, Y)
  return(O)
}


DGS <- generate_data_2
sn = 2
```

```{r child = '2_1_Simu_results_template_1.Rmd'}
```


# Simulation 3
Data structure:  $O = (W, A, Y)$

 * U - exogenous variables
 * W - baseline covariate that is a measure of body condition
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-10 - 3W + 4A + \mathbf{I}(A>2) * 5sin((0.8A)^2 - 2.6)  )]$
  
Outcome of interest: $E_0[Y|a,W]$

```{r check_sys3}

generate_data_3 <- function(n, a=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 1)
  U_A <- rnorm(n, 0, 2)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  Y <- as.numeric(U_Y < plogis(-10 - 3*W + 4*A + 5*sin((0.8*A)^2-2.6)*as.numeric(A > 2)) )
  
  # data frame
  O <- data.frame(W, A, Y)
  return(O)
}


DGS <- generate_data_3
sn = 3
```

```{r child = '2_1_Simu_results_template.Rmd'}
```