---
title: "Simulations of estimating causal effects using undersmoothed HAL"
author: "Seraphina Shi"
date: "2022-09-06"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_lib, include = FALSE, warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(data.table)
library(dplyr)
library(tidyr)
library(foreach)

library(stringr)
library(glmnet)

library(origami)
library(hal9001)
library(tictoc)

library(R.utils)

library(pROC)

library(ggplot2)
library(gridExtra)
library(grid)
```

```{r setup, include = FALSE} 
plotFolder <- here("results","images", "v2")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=plotFolder,
  cache.path=".cache/",
  duplicate.label="allow"
)

source(here("scripts", "scripts_v1", "1_simu_functions_hal9001.R"))
source(here("scripts", "scripts_v1", "1_simu_functions.R"))
```

# Simulation #1
Data structure:  $O = (W, A, Z, Y)$ 

 * U - exogenous variables  
 * W - baseline covariate that is a measure of body condition  
 * A - treatment level based on W, continuous between 0 and 5  
 * Z - intermediate curve based on W and A
 * Y - outcome, indicator of an event ?  
   
 Underlying data generating process, $P_{U,X}$
 
* Exogenous variables:  
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$  
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$  
  + $U_Z \sim Uniform(min = 0, max = 1)$  
  + $U_Y \sim Uniform(min = 0, max = 1)$  
  
* Structural equations F and endogenous variables:  
  + $W =  U_W$  
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$  
  + $Z = \mathbf{I}[U_Z < expit(2-W-A)]$
  + $Y = \mathbf{I}[U_Y < expit(W + 5*A + Z - 0.5 * W * A - 8)]$


```{r}
generate_data_1 <- function(n, a=NA, z=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 1)
  U_A <- rnorm(n, 0, 2)
  U_Z <- runif(n, 0, 1)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  if(is.na(z)){
    Z <- as.numeric(U_Z < plogis(2-W-A))
  } else {
    Z <- rep(z, n)
  }
  
  Y <- as.numeric(U_Y < plogis(W + 5*A + Z - 0.5 * W * A - 8))
  
  # data frame
  O <- data.frame(W, A, Z, Y)
  return(O)
}

obs <- generate_data_1(n=10000)
print(summary(obs))
  
# check positivity violations
cat("Summary of A given W < -1:")
summary(obs$A[obs$W < -1])
cat("Summary of A given -1 < W <= 0:")
summary(obs$A[-1 <= obs$W & obs$W < 0])
cat("Summary of A given 0 < W <= 1:")
summary(obs$A[0 <= obs$W & obs$W < 1])
cat("Summary of A given 1 < W:")
summary(obs$A[1 <= obs$W])

par(mfrow=c(2,3))
plot(obs$W,obs$A)
plot(obs$A,obs$Z)
plot(obs$Z,obs$Y)
plot(obs$W,obs$Z)
plot(obs$W,obs$Y)
plot(obs$A,obs$Y)

glm_fit <- glm(formula = Y ~ W + A + W*A + Z,
               family = binomial,
               data = obs)
summary(glm_fit)
y_preds <- predict(glm_fit, type = "response")
mse <- sum((y_preds - obs$Y)^2)
auc <- auc(obs$Y, y_preds)
print(paste0("    MSE: ", round(mse, 4), ", AUC: ", round(auc, 4)))
```


```{r get_true_psis_1}
# Getting trul value of psi
a_vec <- seq(0,5,0.1)
psi0_a_0 <- c()
psi0_a_1 <- c()

N = 1e+07
data_0_0 <- generate_data_1(n=N, a=0, z=0)
data_0_1 <- generate_data_1(n=N, a=0, z=1)

for (i in 1:length(a_vec)) {
  a <- a_vec[i]

  data_a_0 <- generate_data_1(n=N, a=a, z=0)
  psi0_a_0[i] <- mean(data_a_0$Y - data_0_0$Y)

  data_a_1 <- generate_data_1(n=N, a=a, z=1)
  psi0_a_1[i] <- mean(data_a_1$Y - data_0_1$Y)
}

psi0_pnt <- data.frame(a=rep(a_vec, 2), z=c(rep(1,length(a_vec)), rep(0,length(a_vec))), psi0 = c(psi0_a_1,psi0_a_0))
psi0_10pnt <- psi0_pnt[psi0_pnt$a %in% seq(0.5,5,0.5),]

p <- ggplot() +
    geom_line(data=psi0_pnt %>% mutate(z = as.factor(z)), aes(x=a, y=psi0, color=z, group=z)) + 
    geom_point(data=psi0_10pnt %>% mutate(z = as.factor(z)), aes(x=a, y=psi0, color=z, group=z)) + 
    labs(x="a", y="ATE",
         title = "True Average Treatment Effect \n  P_0(E[Y|a,z] - E[Y|0,z])") +
        theme(plot.title = element_text(hjust = 0.5), #size=8.6, 
              #axis.title.y = element_blank(),
              panel.grid.major.x = element_blank(),
              #axis.text.y = element_text(size=8),
              #axis.title.x = element_text(size=8),
              axis.text = element_text(size=7)) 

p
```



## n = 200
```{r simu_1_n200}
set.seed(123)
n=200

results_200 <- run_simu_1round(generate_data_1, n=200)
psi_10pnts <- merge(as.data.frame(psi0_10pnt), as.data.frame(results_200), by=c("a", "z"))
```


### results
```{r fig.width=8, fig.height=5}
p <- ggplot(data=psi_10pnts %>% mutate(z = as.factor(z)), 
            aes(x=a, y=psi_hat, ymin=CI_lwr, ymax=CI_upr,
                  color=z, group=z),
            alpha = 0.7) +
      geom_line(data=psi0_pnt %>% mutate(z = as.factor(z)), 
                aes(x=a, y=psi0, color=z, group=z),
                alpha = 0.5) +
      geom_errorbar(shape=24, width=0.7) +
        geom_point(data=psi0_10pnt %>% mutate(z = as.factor(z)), 
                 aes(x=a, y=psi0), color = "black") +
        geom_point(shape=17, size=2) +
      labs(x="a", y="ATE",
          title = "Average Treatment Effect \n  E[Y|a,z] - E[Y|0,z]") +
        theme(plot.title = element_text(hjust = 0.5), #size=8.6, 
              #axis.title.y = element_blank(),
              panel.grid.major.x = element_blank(),
              #axis.text.y = element_text(size=8),
              #axis.title.x = element_text(size=8),
              axis.text = element_text(size=7),
              legend.position="none")  +
  facet_grid(~z)

p
```

#### 1000 repetition 
```{r}

```


