---
title: "Simulations of estimating causal effects using undersmoothed HAL"
author: "Seraphina Shi"
date: "2023-03-17"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_lib, include = FALSE, warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(data.table)
library(dplyr)
library(tidyr)
library(foreach)

library(stringr)
library(glmnet)

library(origami)
library(hal9001)
library(tictoc)

library(R.utils)

library(pROC)

library(ggplot2)
library(cowplot)
library(gridExtra)
library(grid)
```

```{r setup, include = FALSE} 
plotFolder <- here("results","images", "v3")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=paste0(plotFolder, "/"),
  cache.path=".cache/",
  duplicate.label="allow"
)

source(here("scripts", "scripts_v3", "1_simu_functions_hal9001.R"))
source(here("scripts", "scripts_v3", "1_simu_functions.R"))
```

# Simulation #1
Data structure:  $O = (W, A, Z, Y)$ 

 * U - exogenous variables  
 * W - baseline covariate that is a measure of body condition  
 * A - treatment level based on W, continuous between 0 and 5  
 * Z - intermediate curve based on W and A
 * Y - outcome, indicator of an event ?  
   
 Underlying data generating process, $P_{U,X}$
 
* Exogenous variables:  
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$  
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$  
  + $U_Z \sim Uniform(min = 0, max = 1)$  
  + $U_Y \sim Uniform(min = 0, max = 1)$  
  
* Structural equations F and endogenous variables:  
  + $W =  U_W$  
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$  
  + $Z = \mathbf{I}[U_Z < expit(2-W-A)]$
  + $Y = \mathbf{I}[U_Y < expit(W + 5*A + Z - 0.5 * W * A - 8)]$


```{r check_sys1}
generate_data_1 <- function(n, a=NA, z=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 1)
  U_A <- rnorm(n, 0, 2)
  U_Z <- runif(n, 0, 1)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  if(is.na(z)){
    Z <- as.numeric(U_Z < plogis(2-W-A))
  } else {
    Z <- rep(z, n)
  }
  
  Y <- as.numeric(U_Y < plogis(W + 5*A + Z - 0.5 * W * A - 8))
  
  # data frame
  O <- data.frame(W, A, Z, Y)
  return(O)
}

obs <- generate_data_1(n=10000)
print(summary(obs))
  
# check positivity violations
cat("Summary of A given W < -1:")
summary(obs$A[obs$W < -1])
cat("Summary of A given -1 < W <= 0:")
summary(obs$A[-1 <= obs$W & obs$W < 0])
cat("Summary of A given 0 < W <= 1:")
summary(obs$A[0 <= obs$W & obs$W < 1])
cat("Summary of A given 1 < W:")
summary(obs$A[1 <= obs$W])

par(mfrow=c(2,4))
hist(obs$W)
hist(obs$A)
plot(obs$W,obs$A)
plot(obs$A,obs$Z)
plot(obs$Z,obs$Y)
plot(obs$W,obs$Z)
plot(obs$W,obs$Y)
plot(obs$A,obs$Y)

glm_fit <- glm(formula = Y ~ W + A + W*A + Z,
               family = binomial,
               data = obs)
summary(glm_fit)
y_preds <- predict(glm_fit, type = "response")
mse <- sum((y_preds - obs$Y)^2)
auc <- auc(obs$Y, y_preds)
print(paste0("    MSE: ", round(mse, 4), ", AUC: ", round(auc, 4)))
```


```{r true_psi_sys1}
# Getting trul value of psi
#------------------------------------------------------------------------------------
# a_vec <- seq(0,5,0.1)
# psi0_a_0 <- c()
# psi0_a_1 <- c()
# 
# N = 1e+07
# data_0_0 <- generate_data_1(n=N, a=0, z=0)
# data_0_1 <- generate_data_1(n=N, a=0, z=1)
# 
# for (i in 1:length(a_vec)) {
#   a <- a_vec[i]
# 
#   data_a_0 <- generate_data_1(n=N, a=a, z=0)
#   psi0_a_0[i] <- mean(data_a_0$Y - data_0_0$Y)
# 
#   data_a_1 <- generate_data_1(n=N, a=a, z=1)
#   psi0_a_1[i] <- mean(data_a_1$Y - data_0_1$Y)
# }
# 
# psi0_pnt <- data.frame(a=rep(a_vec, 2), z=c(rep(1,length(a_vec)), rep(0,length(a_vec))), psi0 = c(psi0_a_1,psi0_a_0))
# psi0_10pnt <- psi0_pnt[psi0_pnt$a %in% seq(0.5,5,0.5),]
# save.image(file=here("data", "rdata", "02_simu_V3_sys1_psi0.RData"))
#------------------------------------------------------------------------------------
load(file=here("data", "rdata", "02_simu_V3_sys1_psi0.RData"))
source(here("scripts", "scripts_v3", "1_simu_functions.R"))
#------------------------------------------------------------------------------------


p <- ggplot() +
    geom_line(data=psi0_pnt %>% mutate(z = as.factor(z)), aes(x=a, y=psi0, color=z, group=z)) + 
    geom_point(data=psi0_10pnt %>% mutate(z = as.factor(z)), aes(x=a, y=psi0, color=z, group=z)) + 
    labs(x="a", y="ATE",
         title = "True Average Treatment Effect \n  P_0(E[Y|a,z] - E[Y|0,z])") +
        theme(plot.title = element_text(hjust = 0.5), #size=8.6, 
              #axis.title.y = element_blank(),
              panel.grid.major.x = element_blank(),
              #axis.text.y = element_text(size=8),
              #axis.title.x = element_text(size=8),
              axis.text = element_text(size=7)) 

p

```



## n = 200
```{r simu_sys1_n200}
nn=200
```


### CV HAL
```{r}
# load(here("data", "rdata", "02_simu_V3_sys1_200_CV.RData"))
# source(here("scripts", "scripts_v3", "1_simu_functions.R"))
```

#### results
```{r simu_sys1_n200_1_cv, fig.width=6, fig.height=4}
set.seed(123)
results_200 <- run_simu_1round(generate_data_1, n=nn)

psi_10pnt <- merge(as.data.frame(psi0_10pnt), as.data.frame(results_200), by=c("a", "z"))

cat(paste0("CV selected lambda (from one sample): ", unique(psi_10pnt$lambda)))


p1 <- plot_perforences_1lambda_alla(psi_10pnt, z_para=1, est_plot_only = T) +
  labs(title = "z=1") + theme(plot.title = element_text(hjust = 0.5))
p0 <- plot_perforences_1lambda_alla(psi_10pnt, z_para=0, est_plot_only = T) +
  labs(title = "z=0") + theme(plot.title = element_text(hjust = 0.5))

p <- grid.arrange(p1, p0, nrow=1,
                  top = textGrob(paste0("Average Treatment Effect,  E[Y|a,z] - E[Y|0,z]",
                                         gp=gpar(fontface = 'bold')))) # gp=gpar(fontsize=11, fontface = 'bold')
```

#### 1000 repetition 

```{r simu_sys1_n200_B_cv, fig.width=6, fig.height=7}
set.seed(123)
simu_results <- run_simu_rep(generate_data_1, n=nn, B=1000, return_all_rslts=T)
cat(paste0("The average of CV selected lambdas (from 1000 sample): ", unique(simu_results$result_summary$lambda)))


cat("z=1:")
p1_list <- plot_perforences_1lambda_alla(simu_results$result_summary, z_para=1, plot_list=F, add_oracal=T)
cat("z=0:")
p0_list <- plot_perforences_1lambda_alla(simu_results$result_summary, z_para=0, plot_list=F, add_oracal=T)

save.image(file=here("data", "rdata", "02_simu_V3_sys1_200_CV.RData"))
```

```{r simu_sys1_n200_B_cv_qq, fig.width=6, fig.height=4}
p1 <- estimation_qqplot(results_list = simu_results$all_results, z_para = 1)
p1
p0 <- estimation_qqplot(results_list = simu_results$all_results, z_para = 0)
p0
```

### Globally Undersmoothed HAL
```{r}
# load(here("data", "rdata", "02_simu_V3_sys1_200_U.RData"))
# source(here("scripts", "scripts_v3", "1_simu_functions.R"))
```

#### results
```{r simu_sys1_n200_1_u, fig.width=6, fig.height=4}
set.seed(123)
n = nn
results_200_under <- run_simu_1round(generate_data_1, n=nn, undersmooth=T)

psi_10pnt <- merge(as.data.frame(psi0_10pnt), as.data.frame(results_200_under), by=c("a", "z"))
cat(paste0("Undersmoothed lambda: ", unique(psi_10pnt$lambda), "\n which is ", unique(psi_10pnt$lambda_scaler), " * lambda_CV"))


p1 <- plot_perforences_1lambda_alla(psi_10pnt, z_para=1, est_plot_only = T) +
  labs(title = "z=1") + theme(plot.title = element_text(hjust = 0.5))
p0 <- plot_perforences_1lambda_alla(psi_10pnt, z_para=0, est_plot_only = T) +
  labs(title = "z=0") + theme(plot.title = element_text(hjust = 0.5))

p <- grid.arrange(p1, p0, nrow=1,
                  top = textGrob(paste0("Average Treatment Effect,  E[Y|a,z] - E[Y|0,z]",
                                         gp=gpar(fontface = 'bold')))) # gp=gpar(fontsize=11, fontface = 'bold')
```

#### 1000 repetition 
```{r simu_sys1_n200_B_u, fig.width=6, fig.height=7}
set.seed(123)

simu_results <- run_simu_rep(generate_data_1, n=nn, B=1000, return_all_rslts=T,  undersmooth=T)


cat(paste0("The average of unsersmoothed lambda (from 1000 sample): ", unique(simu_results$result_summary$lambda),
           "\n which is ", unique(simu_results$result_summary$lambda_scaler), " * the average of 1000 lambda_CV"))

cat("z=1:")
p1_list <- plot_perforences_1lambda_alla(simu_results$result_summary, z_para=1, plot_list=F, add_oracal=T)
cat("z=0:")
p0_list <- plot_perforences_1lambda_alla(simu_results$result_summary, z_para=0, plot_list=F, add_oracal=T)
save.image(file=here("data", "rdata", "02_simu_V3_sys1_200_U.RData"))
```

```{r simu_sys1_n200_B_u_qq, fig.width=6, fig.height=4}
p1 <- estimation_qqplot(results_list = simu_results$all_results, z_para = 1)
p1
p0 <- estimation_qqplot(results_list = simu_results$all_results, z_para = 0)
p0
```

### Oevr a grid of lambda scalers
```{r}
# load(here("data", "rdata", "02_simu_V3_sys1_200_grid.RData"))
# source(here("scripts", "scripts_v3", "1_simu_functions.R"))
```

```{r simu_sys1_n200_B_grid}
set.seed(123)

lambda_scalers <- c(1.2, 1.1, 10^seq(from=0, to=-3, length=30))
# lambda_scalers <- c(1.2, 1.1, seq(from=1, to=0.001, length=28))

simu_results_lists <- list()
for(i in 1:length(lambda_scalers)){
  scaler = lambda_scalers[i]
  simu_results_lists[[i]] <- run_simu_rep(generate_data_1, n=nn, B=1000, lambda_scaler=scaler, return_all_rslts=F,  undersmooth=F)
}
simu_results_all <- do.call("rbind", simu_results_lists) %>% as.data.frame()

save.image(file=here("data", "rdata", "02_simu_V3_sys1_200_grid.RData"))
```


```{r simu_sys1_n200_B_grid_05_1, fig.height=5.5, fig.width=5.5}
# load(here("data", "rdata", "02_simu_V3_sys1_200_grid.RData"))
plot_perforences_alllambda_1a(simu_results_all, a_para = 0.5, z_para = 1, add_oracal=T)
```


```{r simu_sys1_n200_B_grid_15_1, fig.height=5.5, fig.width=5.5}
plot_perforences_alllambda_1a(simu_results_all, a_para = 2.5, z_para = 1, add_oracal=T)
```

```{r simu_sys1_n200_B_grid_5_1, fig.height=5.5, fig.width=5.5}
plot_perforences_alllambda_1a(simu_results_all, a_para = 5, z_para = 1, add_oracal=T)
```

```{r}
# knitr::purl("2_simulations.Rmd")
```


# Simulation 2
```{r check_sys2}
generate_data_2 <- function(n, a=NA, z=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 1)
  U_A <- rnorm(n, 0, 2)
  U_Z <- runif(n, 0, 1)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  if(is.na(z)){
    Z <- as.numeric(U_Z < plogis(2-W-A))
  } else {
    Z <- rep(z, n)
  }
  
  Y <- as.numeric(U_Y < plogis(-4 * W + 5*A + 10*Z + 5*W*Z + 2 * A * Z * W - 15))
  
  # data frame
  O <- data.frame(W, A, Z, Y)
  return(O)
}

obs <- generate_data_2(n=10000)
print(summary(obs))
  
# check positivity violations
cat("Summary of A given W < -1:")
summary(obs$A[obs$W < -1])
cat("Summary of A given -1 < W <= 0:")
summary(obs$A[-1 <= obs$W & obs$W < 0])
cat("Summary of A given 0 < W <= 1:")
summary(obs$A[0 <= obs$W & obs$W < 1])
cat("Summary of A given 1 < W:")
summary(obs$A[1 <= obs$W])


par(mfrow=c(2,4))
hist(obs$W)
hist(obs$A)
plot(obs$W,obs$A)
plot(obs$A,obs$Z)
plot(obs$Z,obs$Y)
plot(obs$W,obs$Z)
plot(obs$W,obs$Y)
plot(obs$A,obs$Y)
```


```{r true_psi_sys2}
# Getting trul value of psi
#------------------------------------------------------------------------------------
a_vec <- seq(0,5,0.1)
psi0_a_0 <- c()
psi0_a_1 <- c()

N = 10000
data_0_0 <- generate_data_2(n=N, a=0, z=0)
data_0_1 <- generate_data_2(n=N, a=0, z=1)

for (i in 1:length(a_vec)) {
  a <- a_vec[i]

  data_a_0 <- generate_data_2(n=N, a=a, z=0)
  psi0_a_0[i] <- mean(data_a_0$Y - data_0_0$Y)

  data_a_1 <- generate_data_2(n=N, a=a, z=1)
  psi0_a_1[i] <- mean(data_a_1$Y - data_0_1$Y)
}

psi0_pnt <- data.frame(a=rep(a_vec, 2), z=c(rep(1,length(a_vec)), rep(0,length(a_vec))), psi0 = c(psi0_a_1,psi0_a_0))
psi0_10pnt <- psi0_pnt[psi0_pnt$a %in% seq(0.5,5,0.5),]

save.image(file=here("data", "rdata", "02_simu_V3_sys2_psi0.RData"))
#------------------------------------------------------------------------------------
# load(file=here("data", "rdata", "02_simu_V3_sys1_psi0.RData"))
# source(here("scripts", "scripts_v3", "1_simu_functions.R"))
#------------------------------------------------------------------------------------


p <- ggplot() +
    geom_line(data=psi0_pnt %>% mutate(z = as.factor(z)), aes(x=a, y=psi0, color=z, group=z)) +
    geom_point(data=psi0_10pnt %>% mutate(z = as.factor(z)), aes(x=a, y=psi0, color=z, group=z)) +
    labs(x="a", y="ATE",
         title = "True Average Treatment Effect \n  P_0(E[Y|a,z] - E[Y|0,z])") +
        theme(plot.title = element_text(hjust = 0.5), #size=8.6,
              #axis.title.y = element_blank(),
              panel.grid.major.x = element_blank(),
              #axis.text.y = element_text(size=8),
              #axis.title.x = element_text(size=8),
              axis.text = element_text(size=7))

p

```


## n = 200
```{r simu_sys2_n200}
nn=200
```


### CV HAL
```{r}
# load(here("data", "rdata", "02_simu_V3_sys1_200_CV.RData"))
# source(here("scripts", "scripts_v3", "1_simu_functions.R"))
```

#### results
```{r simu_sys2_n200_1_cv, fig.width=6, fig.height=4}
set.seed(123)
results_200 <- run_simu_1round(generate_data_2, n=nn)

psi_10pnt <- merge(as.data.frame(psi0_10pnt), as.data.frame(results_200), by=c("a", "z"))

cat(paste0("CV selected lambda (from one sample): ", unique(psi_10pnt$lambda)))


p1 <- plot_perforences_1lambda_alla(psi_10pnt, z_para=1, est_plot_only = T) +
  labs(title = "z=1") + theme(plot.title = element_text(hjust = 0.5))
p0 <- plot_perforences_1lambda_alla(psi_10pnt, z_para=0, est_plot_only = T) +
  labs(title = "z=0") + theme(plot.title = element_text(hjust = 0.5))

p <- grid.arrange(p1, p0, nrow=1,
                  top = textGrob(paste0("Average Treatment Effect,  E[Y|a,z] - E[Y|0,z]",
                                         gp=gpar(fontface = 'bold')))) # gp=gpar(fontsize=11, fontface = 'bold')
```

#### 1000 repetition 

```{r simu_sys2_n200_B_cv, fig.width=6, fig.height=7}
set.seed(123)
simu_results <- run_simu_rep(generate_data_2, n=nn, B=1000, return_all_rslts=T)
cat(paste0("The average of CV selected lambdas (from 1000 sample): ", unique(simu_results$result_summary$lambda)))


cat("z=1:")
p1_list <- plot_perforences_1lambda_alla(simu_results$result_summary, z_para=1, plot_list=F, add_oracal=T)
cat("z=0:")
p0_list <- plot_perforences_1lambda_alla(simu_results$result_summary, z_para=0, plot_list=F, add_oracal=T)

save.image(file=here("data", "rdata", "02_simu_V3_sys1_200_CV.RData"))
```

```{r simu_sys2_n200_B_cv_qq, fig.width=6, fig.height=4}
p1 <- estimation_qqplot(results_list = simu_results$all_results, z_para = 1)
p1
p0 <- estimation_qqplot(results_list = simu_results$all_results, z_para = 0)
p0
```

### Globally Undersmoothed HAL

#### results
```{r simu_sys2_n200_1_u, fig.width=6, fig.height=4}
set.seed(123)
n = nn
results_200_under <- run_simu_1round(generate_data_2, n=nn, undersmooth=T)

psi_10pnt <- merge(as.data.frame(psi0_10pnt), as.data.frame(results_200_under), by=c("a", "z"))
cat(paste0("Undersmoothed lambda: ", unique(psi_10pnt$lambda), "\n which is ", unique(psi_10pnt$lambda_scaler), " * lambda_CV"))


p1 <- plot_perforences_1lambda_alla(psi_10pnt, z_para=1, est_plot_only = T) +
  labs(title = "z=1") + theme(plot.title = element_text(hjust = 0.5))
p0 <- plot_perforences_1lambda_alla(psi_10pnt, z_para=0, est_plot_only = T) +
  labs(title = "z=0") + theme(plot.title = element_text(hjust = 0.5))

p <- grid.arrange(p1, p0, nrow=1,
                  top = textGrob(paste0("Average Treatment Effect,  E[Y|a,z] - E[Y|0,z]",
                                         gp=gpar(fontface = 'bold')))) # gp=gpar(fontsize=11, fontface = 'bold')
```

#### 1000 repetition 
```{r simu_sys2_n200_B_u, fig.width=6, fig.height=7}
set.seed(123)

simu_results <- run_simu_rep(generate_data_2, n=nn, B=1000, return_all_rslts=T,  undersmooth=T)


cat(paste0("The average of unsersmoothed lambda (from 1000 sample): ", unique(simu_results$result_summary$lambda),
           "\n which is ", unique(simu_results$result_summary$lambda_scaler), " * the average of 1000 lambda_CV"))

cat("z=1:")
p1_list <- plot_perforences_1lambda_alla(simu_results$result_summary, z_para=1, plot_list=F, add_oracal=T)
cat("z=0:")
p0_list <- plot_perforences_1lambda_alla(simu_results$result_summary, z_para=0, plot_list=F, add_oracal=T)
save.image(file=here("data", "rdata", "02_simu_V3_sys2_200_U.RData"))
```

```{r simu_sys2_n200_B_u_qq, fig.width=6, fig.height=4}
p1 <- estimation_qqplot(results_list = simu_results$all_results, z_para = 1)
p1
p0 <- estimation_qqplot(results_list = simu_results$all_results, z_para = 0)
p0
```

### Oevr a grid of lambda scalers
```{r}
# load(here("data", "rdata", "02_simu_V3_sys1_200_grid.RData"))
# source(here("scripts", "scripts_v3", "1_simu_functions.R"))
```

```{r simu_sys2_n200_B_grid}
set.seed(123)

lambda_scalers <- c(1.2, 1.1, 10^seq(from=0, to=-3, length=30))
# lambda_scalers <- c(1.2, 1.1, seq(from=1, to=0.001, length=28))

simu_results_lists <- list()
for(i in 1:length(lambda_scalers)){
  scaler = lambda_scalers[i]
  simu_results_lists[[i]] <- run_simu_rep(generate_data_2, n=nn, B=1000, lambda_scaler=scaler, return_all_rslts=F,  undersmooth=F)
}
simu_results_all <- do.call("rbind", simu_results_lists) %>% as.data.frame()

save.image(file=here("data", "rdata", "02_simu_V3_sys2_200_grid.RData"))
```


```{r simu_sys2_n200_B_grid_05_1, fig.height=5.5, fig.width=5.5}
# load(here("data", "rdata", "02_simu_V3_sys1_200_grid.RData"))
plot_perforences_alllambda_1a(simu_results_all, a_para = 0.5, z_para = 1, add_oracal=T)
```


```{r simu_sys2_n200_B_grid_15_1, fig.height=5.5, fig.width=5.5}
plot_perforences_alllambda_1a(simu_results_all, a_para = 2.5, z_para = 1, add_oracal=T)
```

```{r simu_sys2_n200_B_grid_5_1, fig.height=5.5, fig.width=5.5}
plot_perforences_alllambda_1a(simu_results_all, a_para = 5, z_para = 1, add_oracal=T)
```

```{r}
# knitr::purl("2_simulations.Rmd")
```


