---
title: "Simulations of estimating causal effects using undersmoothed HAL"
author: "Seraphina Shi"
date: "2022-09-06"
output: html_document
---

```{r setup, include = FALSE}
plotFolder <- here("results","images")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=plotFolder,
  cache.path=".cache/",
  duplicate.label="allow"
)

set.seed(123)
```

# Simulation #1
Data structure:  $O = (W, A, Z, Y)$ 

 * U - exogenous variables  
 * W - baseline covariate that is a standardized measure of body condition  
 * A - treatment level based on W, continuous between 0 and 10  
 * Z - intermediate curve based on W and A, indicator of taking the treatment  
 * Y - outcome, indicator of an event ?  
   
 Underlying data generating process, $P_{U,X}$
 
* Exogenous variables:  
  + $U_W \sim Uniform(min = 0, max = 1)$  
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$  
  + $U_Z \sim Uniform(min = 0, max = 1)$  
  + $U_Y \sim Uniform(min = 0, max = 1)$  
  
* Structural equations F and endogenous variables:  
  + $W =  U_W$  
  + $A = (10 - 10W + U_A) * \mathbf{I}(0 \leq10 - 10W + U_A \leq 10) + 10* \mathbf{I}(0 \leq10 - 10W + U_A > 10)$  
  + $Z = \mathbf{I}[U_Z < expit(0.2*(W+A))]$
  + $y = \mathbf{I}[U_Y < expit(0.1*W+0.3*A+Z)]$


```{r}
generate_data1 <- function(n){
  # exogenous variables
  U_W <- runif(n, 0, 1)
  U_A <- rnorm(n,0,1)
  U_Z <- runif(n, 0, 1)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  A <- (10 - 10*W + U_A)*as.numeric(10 - 10*W + U_A >= 0 & 10 - 10*W + U_A <= 10) + 10 * as.numeric(10 - 10*W + U_A > 10)
  Z <- as.numeric(U_Z < plogis(0.2*(W+A)))
  Y <- as.numeric(U_Z < plogis(0.1*W + 0.3*A + Z))
  
  # data frame
  O <- data.frame(W, A, Z, Y)
  return(O)
}
```

## n = 1000
```{r}
Obs1 <- generate_data1(1000)
```




