---
title: "Simulations of estimating causal effects using undersmoothed HAL"
author: "Seraphina Shi"
date: "2022-09-06"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_lib, include = FALSE, warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(data.table)
library(dplyr)
library(foreach)

library(stringr)
library(glmnet)

library(origami)
library(hal9001)
library(tictoc)
```

```{r setup, include = FALSE}
plotFolder <- here("results","images")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=plotFolder,
  cache.path=".cache/",
  duplicate.label="allow"
)

source(here("scripts", "1_simu_functions_hal9001.R"))
source(here("scripts", "1_simu_functions.R"))
```

# Simulation #2
Data structure:  $O = (W, A, Z, Y)$ 

 * U - exogenous variables  
 * W - baseline covariate that is a standardized measure of body condition  
 * A - treatment level based on W, continuous between 0 and 10  
 * Z - intermediate curve based on W and A, indicator of taking the treatment  
 * Y - outcome, indicator of an event ?  
   
 Underlying data generating process, $P_{U,X}$
 
* Exogenous variables:  
  + $U_W \sim Uniform(min = 0, max = 1)$  
  + $U_A \sim Normal(\mu=0, \sigma^2 = 4^2)$  
  + $U_Z \sim Uniform(min = 0, max = 1)$  
  + $U_Y \sim Uniform(min = 0, max = 1)$  
  
* Structural equations F and endogenous variables:  
  + $W =  U_W$  
  + $A = bound(10 - 10W + W^2 + U_A, min=0, max=10)$  
  + $Z = \mathbf{I}[U_Z < expit(0.2(W+A))]$
  + $y = \mathbf{I}[U_Y < expit(0.05(W-5A+3*W*A+Z))]$


```{r}
generate_data2 <- function(n, a=NA, z=NA){
  # exogenous variables
  U_W <- runif(n, 0, 1)
  U_A <- rnorm(n,0,4)
  U_Z <- runif(n, 0, 1)
  U_Y <- runif(n, 0, 1)
  
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  10 - 10*W + W*W + U_A
    A[A<=0] = 0
    A[A>=10] = 10
  } else {
    A <- rep(a, n)
  }
  
  if(is.na(z)){
    Z <- as.numeric(U_Z < plogis(0.2*(W+A)))
  } else {
    Z <- rep(z, n)
  }
  
  Y <- as.numeric(U_Z < plogis(0.05 * (-W + 5*A - 3 * W * A - Z)))
  

  # data frame
  O <- data.frame(W, A, Z, Y)
  return(O)
}


obs <- generate_data2(n=10000)
summary(obs)

# check positivity violations
cat("Summary of A given W < 0.25:")
summary(obs$A[obs$W < 0.25])
cat("Summary of A given 0.25 < W <= 0.5:")
summary(obs$A[0.25 <= obs$W & obs$W < 0.5])
cat("Summary of A given 0.5 < W <= 0.75:")
summary(obs$A[0.5 <= obs$W & obs$W < 0.75])
cat("Summary of A given 0.75 < W:")
summary(obs$A[0.75 <= obs$W])

par(mfrow=c(2,3))
plot(obs$W,obs$A)
plot(obs$A,obs$Z)
plot(obs$W,obs$Y)
plot(obs$A,obs$Y)
plot(obs$Z,obs$Y)
```


```{r get_true_psis}
# Getting trul value of psi
a_vec <- seq(0.5,10,0.5)
psi0_a_0 <- c()
psi0_a_1 <- c()

N = 1e+07
data_0_0 <- generate_data2(n=N, a=0, z=0)
data_0_1 <- generate_data2(n=N, a=0, z=1)

for (i in 1:length(a_vec)) {
  a <- a_vec[i]

  data_a_0 <- generate_data2(n=N, a=a, z=0)
  psi0_a_0[i] <- mean(data_a_0$Y - data_0_0$Y)

  data_a_1 <- generate_data2(n=N, a=a, z=1)
  psi0_a_1[i] <- mean(data_a_1$Y - data_0_1$Y)
}
```

## n = 500
```{r simu_n500}
set.seed(123)
n=500
B=1000
# results_500 = run_simu(generate_data = generate_data2, n=n, B=B)
# save.image(file=here("data", "rdata", "02_simulation_2_500.RData"))
```

### results
```{r}
load(here("data", "rdata", "02_simulation_2_500.RData"))

results_500[[1]] <- results_500[[1]] %>% mutate_if(is.numeric, round, digits=4)
results_500[[2]] <- results_500[[2]] %>% mutate_if(is.numeric, round, digits=4)

print(results_500)
```
